Psychological Needs During Intergroup Contact: Three Experience Sampling Studies

### 2022-11-08 Methods-and-Results.tex
M&R shortening

For the core situational need, we asked participants in an open ended
text field:
``\textit{What was your most important goal [during the interaction / this morning / this afternoon]?}''.
Then, with reference to the text entry, we asked how much this core need
was fulfilled during the interaction or the past daytime period:
``\textit{[The interaction / You] fulfilled your goal: [-previous text entry-]}''
on a continuous slider scale ranging from strongly disagree (-50) to
strongly agree (+50). The self-determination theory need measurements
are described in \appref{app:AppendixRobustness}.

@@

We used a multilevel regression to check whether having an interaction
with an outgroup member had a situational (i.e., contemporaneous) effect
within the participants. We find that having an outgroup contact is
indeed associated with significantly more positive outgroup attitudes
(\textit{b} = 2.48, \textit{t}(1,200) = 4.37, \textit{p} \textless{}
.001, \textit{95\%CI}{[}1.37, 3.59{]}), even after controlling for
having an interaction with a non-Dutch (which did not relate to outgroup
attitudes independently). Additionally, while multi-level regressions
are generally robust against unequal cell sizes, we correct for
inequalities by use centered predictors and reintroducing the means as
level two predictors. Nonetheless, we also offer an analysis of the
joint data sets in Online Supplemental Material A (for full results see
\tblref{tab:intergroupGeneralTblLong}, \fgrref{fig:ContactHypothesis},
and Online Supplementary Material
A)\footnote{Interestingly, adding random slopes to this model did not explain additional variance. This is unusual and might indicate that the effect is very consistent across participants. However, the small number of participants, or other measurement issues provide an alternative explanation, which is why we offer a combined data set analyses following the individual studies.}.
Thus, in our first data we find initial evidence that outgroup contacts
show a positive effect on outgroup attitudes within real-life data.
While this offers support for the most basic premise of the intergroup
contact hypothesis (i.e., that any interaction will lead to more
positive outgroup attitudes), this analysis does not consider any notion
of interaction quality. We assess the conditional effects of interaction
numbers and interaction quality ratings after we report on the
individual studies to avoid under-powered analyses.

@@

The main proposal of our article is that the success of an outgroup
contact might be explained by whether or not the contact fulfilled the
person's core situational need. This should, in turn, be due to a higher
perceived contact quality. We sequentially test whether the fulfillment
of the core need during an interaction is (1) related to more positive
outgroup attitudes, (2) higher perceived contact quality, and (3)
whether the variance explained by the core need is subsumed by the
perceived contact quality if considered jointly (for full regression
results see \tblref{tab:intergroupNeedsTblLong} and
\fgrref{fig:AllportNeedFulfillment}).

We find that in the multilevel models, the fulfillment of core
situational needs during outgroup contacts was associated with more
positive outgroup attitudes (random slopes model; \textit{b} = 0.17,
\textit{t}(365) = 2.93, \textit{p} = 0.004, \textit{95\%CI}{[}0.06,
0.29{]}) and also predicted higher perceived contact quality (random
intercept model; \textit{b} = 0.42, \textit{t}(365) = 8.07, \textit{p}
\textless{} .001, \textit{95\%CI}{[}0.32, 0.52{]}). Moreover, when we
consider the influences of core need fulfillment and contact quality on
outgroup attitudes jointly, we find that virtually all variance is
explained by perceived contact quality (random slopes mode; \textit{b} =
0.23, \textit{t}(364) = 4.22, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.12, 0.33{]}) and only little unique variance is
still explained by core need fulfillment (\textit{b} = 0.04,
\textit{t}(364) = 0.64, \textit{p} = 0.523, \textit{95\%CI}{[}-0.08,
0.15{]}). We thus find support for our hypotheses and can conclude that
in this data set the fulfillment of core situational needs had a
significant influence on outgroup attitudes. Additionally, this effect
is likely explained by its effect through perceived contact quality.

@@

As in Study 1, attitudes towards the Dutch majority outgroup were again
measured using the feeling thermometer. All survey details are also
available in our OSF repository \citep{KreienkampMasked2022a} and Online
Supplemental Materials A.

@@

We tested the most general contact hypothesis as we did for Study 1. We,
again, used a multilevel regression analysis to check whether having an
interaction with an outgroup member had a situational (i.e.,
contemporaneous) effect within the participants. We find that having an
outgroup interaction is indeed associated with significantly more
positive outgroup attitudes within the participants (random slopes
model; \textit{b} = 2.83, \textit{t}(4,850) = 3.57, \textit{p}
\textless{} .001, \textit{95\%CI}{[}1.28, 4.38{]}), even after
controlling for having an interaction with a non-Dutch person (which did
not relate to outgroup attitudes independently). We again added the
participant means back into the model. Because these are the means of
the contact dummy variables they are the proportion of measurements that
included a particular contact. This not only adjusts the model for
unequal cell sizes but allows us to see whether between subject
differences in the relative number of interactions had an effect on
outgroup attitudes. We find that in this data set participant-level
outgroup contact proportions were also a positive predictor of outgroup
attitudes (\textit{b} = 26.55, \textit{t}(110) = 2.90, \textit{p} =
0.004, \textit{95\%CI}{[}8.61, 44.46{]}). The relative number of
non-outgroup interactions showed no such effect (for full results see
\tblref{tab:intergroupGeneralTblLong}, \fgrref{fig:ContactHypothesis},
as well as Online Supplementary Material A). Thus, in our second data
set we find that outgroup contacts show a positive effect on outgroup
attitudes in the moment and on average between participants.

@@

Attitudes towards the Dutch majority outgroup was again measured using
the feeling thermometer, as in studies one and two. All survey details
are also available in our OSF repository \citep{KreienkampMasked2022a}
and Online Supplemental Materials A.

@@

Additionally, in a multilevel regression to check for contemporaneous
within person effects, we found that having an outgroup interaction was
indeed associated with significantly more positive outgroup attitudes
within the participants (random slopes model; \textit{b} = 5.57,
\textit{t}(3,834) = 6.52, \textit{p} \textless{} .001,
\textit{95\%CI}{[}3.90, 7.23{]}), even after controlling for having an
interaction with a non-Dutch (which did not relate to outgroup attitudes
independently; for full results see
\tblref{tab:intergroupGeneralTblLong}, \fgrref{fig:ContactHypothesis},
and Online Supplementary Material A). Thus, in our third data set we
found within person contemporaneous effect of intergroup contact was
consistent across all three studies.

@@

Similarly to the previous studies, we again sequentially tested whether
the fulfillment of the core need during an interaction was (1) related
to more positive outgroup attitudes, (2) higher perceived interaction
quality, and (3) whether the variance explained by the core need is
assumed by the perceived interaction quality if considered jointly. We
find that the fulfillment of the core need during outgroup contacts was
associated with more positive outgroup attitudes (random slopes model;
\textit{b} = 0.19, \textit{t}(1,601) = 5.29, \textit{p} \textless{}
.001, \textit{95\%CI}{[}0.12, 0.27{]}) and also predicted higher
perceived interaction quality (random slopes model; \textit{b} = 0.43,
\textit{t}(1,605) = 8.35, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.33, 0.54{]}). Additionally, once we consider the
influences of core need fulfillment and interaction quality on outgroup
attitudes jointly, we find that perceived interaction quality is a
substantially stronger predictor (random slopes mode; \textit{b} = 0.17,
\textit{t}(1,600) = 6.86, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.12, 0.22{]}) and the unique variance explained by
core need fulfillment was roughly half of its original effect size
(\textit{b} = 0.11, \textit{t}(1,600) = 3.50, \textit{p} \textless{}
.001, \textit{95\%CI}{[}0.05, 0.17{]}; for full results see
\tblref{tab:intergroupNeedsTblLong} and
\fgrref{fig:AllportNeedFulfillment}). These results indicate that in
this data set outgroup attitudes were significantly predicted by the
fulfillment of core situational needs and this effect partially emerged
via its effect on perceived interaction quality.

@@

To test wether Allport's conditions or the core need fulfillment are
better at predicting outgroup attitudes, we first assess relative model
performance indices (i.e., Akaike information criterion, and Bayesian
information criterion), and then consider the two predictors in a joint
model to see whether the two approaches predict the same variance in
outgroup attitudes. When comparing the model selection indices, we find
that the fulfillment of the situational core need indeed performs
slightly better than the model using Allport's conditions
(\(AIC_{CoreNeed}\) 12632.02 \textless{} 12651.59 \(AIC_{Allport}\), and
\(BIC_{CoreNeed}\) 12664.55 \textless{} 12684.12 \(BIC_{Allport}\)).
Additionally, when considering the predictors jointly, we find that both
significantly predict outgroup attitudes with similar sized regression
parameters (random slopes model; Allport's Conditions: \textit{b} =
0.16, \textit{t}(1,600) = 4.92, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.09, 0.24{]}, Core Need: \textit{b} = 0.14,
\textit{t}(1,600) = 3.85, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.08, 0.17{]}; also see
\tblref{tab:intergroupNeedsTblLong} and
\fgrref{fig:AllportNeedFulfillment}). This indicates that, although both
Allport's conditions and the core need fulfillment seem to (in part)
work through perceived interaction quality, they explain different
aspects of the variance in outgroup attitudes and do not constitute one
another.

\section{Embeddedness and Stability across Studies}

@@

We begin by assessing the stability of our main analyses across the
three studies. To do so, we plot the effect sizes of each parameter of
interest in a forest plot. This affords us a chance to spot stable as
well as unreliable effects more easily. While not the focus of our
efforts here, forest plots are commonly accompanied with a meta-analytic
effect size that summarizes the effect size across studies.

We assess forest plots of the basic contact hypothesis test
(\fgrref{fig:ContactHypothesis}), as well as the core need fulfillment
proposal (\fgrref{fig:AllportNeedFulfillment}). For the basic contact
hypothesis test we see that outgroup contact has a strong and consistent
effect on outgroup attitudes while interactions with non-outgroup
members consistently have no meaningful effect on outgroup attitudes.
While this would be expected from the general intergroup contact
literature, this is not a trivial finding. We are among the first to
assess the contact hypothesis using real-life intensive longitudinal
data and these close to event assessments are comparable but not
equivalent to the assessments found in cross-sectional data. When
looking at the fulfillment of core needs during intergroup contacts, we
find that the motivational mechanism is consistently a meaningful
predictor of outgroup attitudes and interaction quality perceptions (see
\fgrref{fig:AllportNeedFulfillment} A and B). We also see that the
effect of core need fulfillment on outgroup attitudes is strongly
reduced when modeled together with interaction quality perceptions,
supporting our assertion that interaction quality offers a psychological
mechanism for the direct effect.

@@

During the main analyses, we have thus far shown that participants held
more positive outgroup attitudes following intergroup contacts and that
a perceived interaction quality was associated with more positive
outgroup attitudes following an intergroup contact. We have, however,
not brought the two elements of the contact hypothesis together in a
single analysis. This is mainly due to the difficulty that interaction
quality ratings are only possible when participants had an interaction
--- which means that the two factors cannot be assessed in a multilevel
analysis that models individual measurement occasions. This has not been
an issue in the past literature because experimental and cross-sectional
data collections have only considered cases in which intergroup contacts
were present. Outside of the lab, the most common method has been to ask
participants to recall how many interactions they had over the past
month, how positive these interactions were, and how positive they feel
towards the outgroup in general. The effects of intergroup contacts,
interaction quality, and outgroup attitudes are thus summarized by the
participant in a cognitive recall aggregation. Intensive longitudinal
data also allows for aggregation of all three variables, with the
differences that the measurements are aggregated from individual reports
that are close to the actual experience. However, this aggregation
process summarizes the dense multitude of measurement occasions as
single summary statistics per participants. This, by extension, means
that we face statistical power restrictions on the person level and have
thus decided to only present this analysis based on a combined data set
of all three studies.

We run a linear regression model where average outgroup attitudes are
predicted by the number of interactions and the average interaction
quality ratings of all participants across the three studies. We do so
while controlling for the possible effects of study-specific
differences. To include the study-membership as a control variable, we
use the student sample (Study 2) as the reference group because it is
both the largest and the most homogeneous study. Looking at the overall
model, we find that the model predicts a statistically significant
amount of variance in average outgroup attitudes (\textit{F}(9, 189) =
3.39, \textit{p} = 0.001, \textit{R\textsuperscript{2}} = 0.14). Looking
at the individual effects, we find that only the number of outgroup
interactions has a clear association with average outgroup attitudes
(\textit{b} = 0.55, \textit{t}(189) = 2.82, \textit{p} = 0.005,
\textit{95\%CI}{[}0.16, 0.93{]}). The average interaction quality
perceptions had a much smaller effect (\textit{b} = 0.29,
\textit{t}(189) = 2.09, \textit{p} = 0.038, \textit{95\%CI}{[}0.02,
0.56{]}), and we do not find the most important interaction effect at
all. In short, the effect of outgroup interactions did not depend on the
average interaction quality, so that participants who on average had
more positive interactions were not statistically more likely to show
higher average outgroup attitudes than those with more negative
interactions. This result is surprising because it contradicts the tenet
of the intergroup contact hypothesis that intergroup contact depends on
positive contact.

@@

In terms of the content of the topics, we find that a number of topics
are primarily task-oriented, where participants hope to increase their
study, research, presentation, or work performance. Opposing the task
and work oriented needs, are a wide variety of leisure-related wishes.
Specifically, relaxation and entertainment wishes were prominent terms
across several topics. Additionally, a number of clusters is primarily
relationship-oriented, so that participants sought contact with outgroup
members for intimate and casual social contact in itself. Similarly,
socializing and celebrations were also explicit social needs (incl.,
parties). Another set of topics was more practically oriented, when
participants sought to share food and cook cook together, or had
organizational needs (e.g., housing, cleaning, and living).

A reasonable set of contact goals were specifically migration-specific
(e.g., wish to learn about culture, politics, and language) or were
concerned inquiry and information needs more generally (e.g., seeking
answers, bureaucratic information). A further set of topics was
specifically geared towards a wish to experience cultural products
(e.g., music, theater, food) or had travel related goals in their
interactions with the majority group members.

One interesting observation was the importance of contact goals specific
to contact through the medical and public health system. This goal type
was partly specific to our young medical professionals sample (e.g.,
working with patients, treatment), but also more broadly interactions of
newcomers with the outgroup majority as patients themselves (e.g.,
therapeutic goals). Also health, fitness, and personal improvement goals
(e.g., sports and music) were common goals that participants shared
during the interactions with the majority group members. A final topic
was that of spiritual, religious, and otherwise transcendental needs
(incl.~meditation, prayer, religious services).

Thus, almost all extracted topics fall into broader or narrower need
concepts that are commonly discussed within the psychological needs
literature \citep[e.g.,][]{Orehek2018a} and offer insight into a core
aspect of the migration experience that has remained broadly
under-explored \citep[][]{Kreienkamp2022d}.

>>>>

For the core situational need, we asked participants in an open-ended
text field:
``\textit{What was your most important goal [during the interaction / this morning / this afternoon]?}''.
Then, with reference to the text entry, we asked how much this core need
was fulfilled during the interaction or the past daytime period:
``\textit{[The interaction / You] fulfilled your goal: [-previous text entry-]}''
on a continuous slider scale ranging from strongly disagree (-50) to
strongly agree (+50). The self-determination theory need measurements
are described in \appref{app:AppendixRobustness}.

@@

We used a multilevel regression to check whether having an interaction
with an outgroup member had a situational (i.e., contemporaneous) effect
within the participants. We find that having an outgroup contact is
indeed associated with significantly more positive outgroup attitudes
(\textit{b} = 2.48, \textit{t}(1,200) = 4.37, \textit{p} \textless{}
.001, \textit{95\%CI}{[}1.37, 3.59{]}), even after controlling for
having an interaction with a non-Dutch (which did not relate to outgroup
attitudes independently). Additionally, while multi-level regressions
are generally robust against unequal cell sizes, we correct for
inequalities by use centered predictors and reintroducing the means as
level two predictors (\citealp{Yaremych2021}; for full results see
\tblref{tab:intergroupGeneralTblLong}, \fgrref{fig:ContactHypothesis},
and Online Supplementary Material
A)\footnote{Interestingly, adding random slopes to this model did not explain additional variance. This is unusual and might indicate that the effect is very consistent across participants. However, the small number of participants, or other measurement issues provide an alternative explanation, which is why we offer a combined data set analyses following the individual studies.}.
Thus, in our first data we find initial evidence that outgroup contacts
show a positive effect on outgroup attitudes within real-life data.

@@

The main proposal of our article is that the success of an outgroup
contact might be explained by whether or not the contact fulfilled the
person's core situational need. This should, in turn, be due to a higher
perceived contact quality. We sequentially test whether the fulfillment
of the core need during an interaction is (1) related to more positive
outgroup attitudes, (2) higher perceived contact quality, and (3)
whether the variance explained by the core need is subsumed by the
perceived contact quality if considered jointly. We find that in the
multilevel models, the fulfillment of core situational needs during
outgroup contacts was associated with more positive outgroup attitudes
(random slopes model; \textit{b} = 0.17, \textit{t}(365) = 2.93,
\textit{p} = 0.004, \textit{95\%CI}{[}0.06, 0.29{]}) and also predicted
higher perceived contact quality (random intercept model; \textit{b} =
0.42, \textit{t}(365) = 8.07, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.32, 0.52{]}). Moreover, when we consider the
influences of core need fulfillment and contact quality on outgroup
attitudes jointly, we find that virtually all variance is explained by
perceived contact quality (random slopes mode; \textit{b} = 0.23,
\textit{t}(364) = 4.22, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.12, 0.33{]}) and only little unique variance is
still explained by core need fulfillment (\textit{b} = 0.04,
\textit{t}(364) = 0.64, \textit{p} = 0.523, \textit{95\%CI}{[}-0.08,
0.15{]}). We thus find support for our hypotheses and can conclude that
in this data set the fulfillment of core situational needs had a
significant influence on outgroup attitudes. Additionally, this effect
is likely explained by its effect through perceived contact quality.

@@

As in Study 1, attitudes towards the Dutch majority outgroup were again
measured using the feeling thermometer.

@@

We tested the most general contact hypothesis as we did for Study 1. We
find that having an outgroup interaction is indeed associated with
significantly more positive outgroup attitudes within the participants
(random slopes model; \textit{b} = 2.83, \textit{t}(4,850) = 3.57,
\textit{p} \textless{} .001, \textit{95\%CI}{[}1.28, 4.38{]}), even
after controlling for having an interaction with a non-Dutch person
(which did not relate to outgroup attitudes independently). We again
added the participant means back into the model. We find that in this
data set participant-level outgroup contact proportions were also a
positive predictor of outgroup attitudes (\textit{b} = 26.55,
\textit{t}(110) = 2.90, \textit{p} = 0.004, \textit{95\%CI}{[}8.61,
44.46{]}). The relative number of non-outgroup interactions showed no
such effect (for full results see \tblref{tab:intergroupGeneralTblLong},
\fgrref{fig:ContactHypothesis}, as well as Online Supplementary Material
A). Thus, in our second data set we also find that outgroup contacts
show a positive effect on outgroup attitudes in the moment. We
additionally also find an average between participant effect of the
relative number of interactions participants had.

@@

Attitudes towards the Dutch majority outgroup was again measured using
the feeling thermometer, as in studies one and two.

@@

In the multilevel regression, we find that having an outgroup
interaction was again associated with significantly more positive
outgroup attitudes within the participants (random slopes model;
\textit{b} = 5.57, \textit{t}(3,834) = 6.52, \textit{p} \textless{}
.001, \textit{95\%CI}{[}3.90, 7.23{]}), even after controlling for
having an interaction with a non-Dutch (which did not relate to outgroup
attitudes independently; for full results see
\tblref{tab:intergroupGeneralTblLong}, \fgrref{fig:ContactHypothesis},
and Online Supplementary Material A). Thus, in our third data set we
find that the within-person contemporaneous effect of intergroup contact
was consistent across all three studies.

@@

Analogous to the previous studies, we again sequentially tested whether
the fulfillment of the core need during an outgroup interaction was (1)
related to more positive outgroup attitudes, (2) higher perceived
interaction quality, and (3) whether the variance explained by the core
need is assumed by the perceived interaction quality if considered
jointly. We find that the fulfillment of the core need during outgroup
contacts was associated with more positive outgroup attitudes (random
slopes model; \textit{b} = 0.19, \textit{t}(1,601) = 5.29, \textit{p}
\textless{} .001, \textit{95\%CI}{[}0.12, 0.27{]}) and also predicted
higher perceived interaction quality (random slopes model; \textit{b} =
0.43, \textit{t}(1,605) = 8.35, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.33, 0.54{]}). Additionally, once we consider the
influences of core need fulfillment and interaction quality on outgroup
attitudes jointly, we find that perceived interaction quality is a
substantially stronger predictor (random slopes mode; \textit{b} = 0.17,
\textit{t}(1,600) = 6.86, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.12, 0.22{]}) and the unique variance explained by
core need fulfillment was roughly half of its original effect size
(\textit{b} = 0.11, \textit{t}(1,600) = 3.50, \textit{p} \textless{}
.001, \textit{95\%CI}{[}0.05, 0.17{]}; for full results see
\tblref{tab:intergroupNeedsTblLong} and
\fgrref{fig:AllportNeedFulfillment}). As with the previous two studies,
these results indicate that in this data set outgroup attitudes were
significantly predicted by the fulfillment of core situational needs and
the results suggest that this might be due to higher perceived
interaction quality.

@@

To test whether Allport's conditions or the core need fulfillment are
better at predicting outgroup attitudes, we first assess relative model
performance indices (i.e., Akaike information criterion, and Bayesian
information criterion), and then consider the two predictors in a joint
model to see whether the two approaches predict the same variance in
outgroup attitudes. When comparing the model selection indices, we find
that the fulfillment of the situational core need indeed performs
slightly better than the model using Allport's conditions
(\(AIC_{CoreNeed}\) 12632.02 \textless{} 12651.59 \(AIC_{Allport}\), and
\(BIC_{CoreNeed}\) 12664.55 \textless{} 12684.12 \(BIC_{Allport}\)).
Additionally, when considering the predictors jointly, we find that both
significantly predict outgroup attitudes with similar sized regression
parameters (random slopes model; Allport's Conditions: \textit{b} =
0.16, \textit{t}(1,600) = 4.92, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.09, 0.24{]}, Core Need: \textit{b} = 0.14,
\textit{t}(1,600) = 3.85, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.08, 0.17{]}; also see
\tblref{tab:intergroupNeedsTblLong} and
\fgrref{fig:AllportNeedFulfillment}). This indicates that, although both
Allport's conditions and the core need fulfillment seem to (in part)
work through perceived interaction quality, they explain different
aspects of the variance in outgroup attitudes and do not constitute one
another.

\section{Robustness, Stability, and Embeddeness across Studies}

@@

We begin by assessing the stability of our main analyses across the
three studies. To do so, we plot the effect sizes of each parameter of
interest in a forest plot. This affords us a chance to spot stable as
well as unreliable effects more easily. While not the focus of our
efforts here, forest plots are commonly accompanied with a meta-analytic
effect size that summarizes the effect size across studies. For the
basic contact hypothesis test we see that outgroup contact has a strong
and consistent effect on outgroup attitudes while interactions with
non-outgroup members consistently have no meaningful effect on outgroup
attitudes (\fgrref{fig:ContactHypothesis}). While this would be expected
from the general intergroup contact literature, this is not a trivial
finding. We are among the first to assess the contact hypothesis using
real-life intensive longitudinal data and these close to event
assessments are comparable but not equivalent to the assessments found
in cross-sectional data. When looking at the fulfillment of core needs
during intergroup contacts, we find that the motivational mechanism is
consistently a meaningful predictor of outgroup attitudes and
interaction quality perceptions (see
\fgrref[ A and B]{fig:AllportNeedFulfillment}). We also see that the
effect of core need fulfillment on outgroup attitudes is strongly
reduced when modeled together with interaction quality perceptions,
supporting our assertion that interaction quality offers a psychological
mechanism for the direct effect.

@@

During the main analyses, we have thus far shown that participants held
more positive outgroup attitudes following intergroup contacts and that
a perceived interaction quality was associated with more positive
outgroup attitudes following an intergroup contact. We have, however,
not brought the two elements of the contact hypothesis together in a
single analysis. To jointly test the effects of contact frequency and
average interaction quality, we run a linear regression model where
average outgroup attitudes are predicted by the number of interactions
and the average interaction quality ratings of all participants across
the three studies. We do so while controlling for the possible effects
of study-specific differences. To include the study-membership as a
control variable, we use the student sample (Study 2) as the reference
group because it is both the largest and the most homogeneous study.
Looking at the overall model, we find that the model predicts a
statistically significant amount of variance in average outgroup
attitudes (\textit{F}(9, 189) = 3.39, \textit{p} = 0.001,
\textit{R\textsuperscript{2}} = 0.14). Looking at the individual
effects, we find that only the number of outgroup interactions has a
clear association with average outgroup attitudes (\textit{b} = 0.55,
\textit{t}(189) = 2.82, \textit{p} = 0.005, \textit{95\%CI}{[}0.16,
0.93{]}). The average interaction quality perceptions had a much smaller
effect (\textit{b} = 0.29, \textit{t}(189) = 2.09, \textit{p} = 0.038,
\textit{95\%CI}{[}0.02, 0.56{]}), and we do not find the most important
interaction effect at all. In short, the effect of interaction frequency
did not depend on the average interaction quality.

@@

In terms of the content of the topics, we find that a number of topics
are primarily task-oriented, where participants hope to increase their
study, research, presentation, or work performance. Opposing the task
and work oriented needs, are a wide variety of leisure-related wishes.
Specifically, relaxation and entertainment wishes were prominent terms
across several topics. Additionally, a number of clusters is primarily
relationship-oriented, so that participants sought contact with outgroup
members for intimate and casual social contact in itself. Similarly,
socializing and celebrations were also explicit social needs (incl.,
parties). This also included a subtopic of social spiritual, religious,
and otherwise transcendental needs (incl.~meditation, prayer, religious
services). Among the leisure-oriented topics was also a set of contact
goals that were specifically migration-specific (e.g., wish to learn
about culture, politics, and language) or were concerned inquiry and
information needs more generally (e.g., seeking answers, bureaucratic
information). A similar set of topics was specifically geared towards a
wish to experience cultural products (e.g., music, theater, food) or had
travel related goals in their interactions with the majority group
members.

One interesting observation was the importance of contact goals specific
to contact through the medical and public health system. This goal type
was partly specific to our young medical professionals sample (e.g.,
working with patients, treatment), but also more broadly interactions of
newcomers with the outgroup majority as patients themselves (e.g.,
therapeutic goals). Also health, fitness, and personal improvement goals
(e.g., sports and music) were common goals that participants shared
during the interactions with the majority group members. Another topic
less commonly found within the literature were more practically oriented
needs, where participants sought to share food and cook together, or had
organizational needs (e.g., housing, cleaning, and co-living). In sum,
almost all extracted topics fall into broader or narrower need concepts
that are commonly discussed within the psychological needs literature
\citep[e.g.,][]{Orehek2018a} and offer insight into a core aspect of the
migration experience that has remained broadly under-explored
\citep[][]{Kreienkamp2022d}\footnote{It should be noted that these topic models are not without limitations, especially because they depend on a small set of hyper parameters that determine the characteristics of the embedding, dimension reduction, clustering, and core term extraction. The authors have have also inspected a major subset of the free-text responses manually and are under the impression that the topics described here accurately represent the broader content.}.
